---
title: "hw6-2"
author: "juan"
date: '2017-11-06'
output: github_document
---
```{r}
library(gapminder)
library(purrr)
library(devtools)
library(stringr)
library(listviewer)
library(tibble)
library(forcats)
library(ggplot2)
library(png)
library(grid)
library(RColorBrewer)
library(gapminder)
library(readxl)
suppressPackageStartupMessages(library(dplyr))
suppressMessages(library(gridExtra) )
suppressMessages(library(mgcv))
suppressMessages(library(jsonlite))
suppressPackageStartupMessages(library(tidyverse))
```

### Welcome to part two of  HMW6
stumbling from step 1 :(
The command: install.github(jennybc/candy)  did not work for some reason. So I went ahead and forked the candy repo and tried to install it from my repo. Failed again, so I had to download the .xlsx to my mac and import the dataset manually using this code:
```{r}
library(readxl)
suppressMessages(candy <- read_excel("~/Downloads/CANDY-HIERARCHY-2015-SURVEY-Responses.xlsx"))
```

#### Intial thoughts and observations:
In all my years as data scientist **(0 years)**, I have learnt that nothing with data is easy. Add to that the fact that my r skills aren't sharp enought yet (getting there tho) and that I'm an expert in getting errors in Rstutio. So, yes, it was all confusing from the begining.   
Anyway, this analysis will be based on the sentiments expressed by participants of the halloween surver 2015 and will serve the purpose of revealing what is people's favourite candy and most hated candy. In order to do this, many issue need to be solved first, for example, I should clean a bit of things:
* Irrelevant Data: leave out all non-candy products in the dataset
* Format of the names in the columns:the name of the columns are messy, lists[] and brackers()
* Coercing datatypes: different datatypes in some attributes (age appears as characters instead of numeric, sign of something fishy)
* Dealing with missing values:I have not learnt how to deal with NA, so i'll replace them for "None" or ignore them.
* String Formats: lowercase all values
 I'll drop some columns that I find irrelevant. "spotted dicks", really? what is that? swedish fish? is there a fish flavour candy?? yuck!! All the columns that have opinions will be ignored because time is scarce and that takes a lot of coding, maybe in future assignments it could be taken into consideration, cleaned and analysed. I'll keep only candy and columns with joy and despair. 
```{r}
#summary(candy)
#str(candy)
#glimpse(candy)
#class(candy)
#head(candy, n=20)
```
 
* I did do the exploratory using the above functions part but I will not execute the code because the results take many pages. Once I clean the data I'll print them. The number of candy is very broad, in my country there are 8 kinds of candy tops, I don't know many of the ones listed here.
* Since there is no attribute that connect the data to "real candy" and "not candy", I'll select the true candy manually and I'll leave out the non-candy.
```{r}
preselectedcandy<- c( "How old are you?","[Butterfinger]", "[100 Grand Bar]", "[Bonkers]","[Bubble Gum]", "[Caramellos]", "[Snickers]", "[Gummy Bears straight up]", "[Mars]", "[Maynards]", "[Skittles]", "[Kit Kat]","[Twix]", "[Starburst]", "[Nerds]", "[Nestle Crunch]", "[Milky Way]", "[Reese’s Peanut Butter Cups]", "[Lollipops]", "[Hershey’s Milk Chocolate]", "[Regular M&Ms]") 
data<-candy %>% select(preselectedcandy)
```


* The names contain square brackets which is unnecessary and some of them have extra information such as "regular" or "straight up". I'll clean the names first using stringr::str_replace_all(), str_trim() and some regex. (wooohoo already profiting from knowledge gained in part one of the hmwk)
```{r}
formated_names<-str_replace_all(preselectedcandy,c("\\[" = " ", "\\]" = " ")) %>% str_trim()
names(data)<- formated_names
```


```{r}
new_names<-names(data) %>% str_replace_all(c("How old are you\\?"="Age", "Gummy Bears straight up"= "Gummy Bears", "Regular M&Ms"= "M&Ms")) %>% tolower()
names(data)<- new_names
data %>% head()
```

* Time to tackle the age format. It appears as chr instead of numeric. It is because down there some values are strings ("old enough", "really old", "none of your business"...). Since one column can only have one datatype, it came as character. I'll change it into numeric which will make coerce characers into NAs. I'll also try to find outliers.
```{r hist}
data$age<-as.integer(data$age) 
hist(data$age)
```

* It is normal at this stage to see such wacky graphs. Clearly there are some mistakes with the data. We'll examine the tables to see age clearly. All values are now in integer format, so strings were coerced into NA.
```{r}
data %>% arrange(desc(age)) %>% head()
```

* UUUhhh there were some reaaally old folks there, where do they find the joy? After a 120 years old it's all despair.
Possible explanations for these numbers is data entry typos (example 2000 instead of 20.00, or some people writing the years and months as well 12.3). Since there is no date of birth, I have no way to get the real age of the outliers, so I'll discart people over 100 years old, 0 years old and NA values in the age column.

```{r}
data2<- data %>% filter(age<100 & age!= 0 & age != "NA")%>% arrange(age)
data2 %>% head(10)
data2 %>% tail(10)
```

* We can double check our work and see NA values gone.
```{r}
summary(data2)
```

Next step, replace NA values in other columns as "neutral". Why don't I just ditch them? because it is normal not to have feelings for every single brand of candy. Not having feelings for a brand of candy is also valuable information for the analysis since it may indicates candy types that don't sell very well.  
```{r}
data_clean <- as.data.frame(data2)
data_clean[is.na(data_clean)] <- "NEUTRAL"
data_clean %>% head()

```

Numbers look more consistent and realistic than before. Let Lowercase the rest of the data.
```{r}
dflong<- data_clean  %>% gather(brand , emotion,-age) %>% mutate(emotion= tolower(emotion))
dfcount<-dflong %>%  group_by(brand, emotion, age) %>% count(emotion) %>% arrange(brand, desc(n)) 
write.csv(dfcount, "candyc.csv") 
head(dfcount)
```

* Now that the data is manageable we can do some analysis. I'll include some generation labels to group the ages. Based on this grouping, I"ll examine the favourite candy per generation and the differences in emotions in all groups and candy types.

```{r}
centennial_ages<- 1:21
millennial_ages<- 22:40
generationX_ages<- 41:52
baby_Boomers_ages<- 53:71
traditionalists_ages<- 72:100
```

* Reshaping data to make it easier to use.  I use the age values to group people into generations.
```{r}
centennials<-dfcount %>% filter(age %in% centennial_ages) 
centennials %>% head()

```


#### OK, let's plot!! 
* I could not invest much time polishing all the features of the graphs. I did use different kinds of plots tho.
###Kids are not big fans of bonkers and maynards compare to other brands.
```{r centennial_ages}
cent_emotions %>% ggplot(aes( x = brand, y = n, fill= emotion)) + 
  geom_bar(stat= "identity", position = "stack") +
    theme(axis.text.x = element_text(angle=90))+ ggtitle("Emotions: Centennials")
```


* This one needs more work in order to improve readability. I know colors are not the best, something to improve in the future with more time.

```{r }
millennials<-dfcount %>% filter(age %in% millennial_ages) 


```


* Millenials like Reese's peanut butter cups, kit kats and skittles a lot. 
```{r millennial_ages}
mill_emotions %>% ggplot(aes(x = emotion, y = n)) + geom_bar(aes(fill= brand),stat= "identity", position = "dodge") +
    theme(axis.text.x = element_text(angle=90))+ ggtitle("Emotions: Millenials")
```


* Reese's peanut butter cups, kit kats and snickers seem to do well among GenXers too.
```{r generationX_ages}
genx<-dfcount %>% filter(age %in% generationX_ages) 
genx %>% head(10)
genx_emotions<- genx %>% group_by(brand, emotion) %>%  count(emotion) %>% arrange(brand, desc(n))
genx_emotions %>% ggplot(aes(x = emotion, y = brand, color= n)) + geom_point(aes(size= n))+ ggtitle("Emotions: GenerationXers")
```

* People, no matter the age, like bonkers, bubble gum or maynards very much. 

```{r}
baby_b<-dfcount %>% filter(age %in% baby_Boomers_ages) 
baby_b %>% head(10)
baby_b_emotions<- baby_b %>% group_by(brand, emotion) %>%  count(emotion) %>% arrange(brand, desc(n))
baby_b_emotions %>% ggplot(aes(y = n, x = brand, color= emotion)) + geom_point(aes(size= n)) +
    theme(axis.text.x = element_text(angle=90))+ ggtitle("Emotions: Millenials")+ ggtitle("Emotions: Baby Boomers")
```

* Butterscotch would have killed in here but sadly is not in the list. 
```{r traditionalists_ages}
traditionalists<-dfcount %>% filter(age %in% traditionalists_ages) 
traditionalists %>% head(10)
traditionalists_emotions<- baby_b %>% group_by(brand, emotion) %>%  count(emotion) %>% arrange(brand, desc(n))
traditionalists_emotions %>% ggplot(aes(x = n, y = brand, color= emotion)) + geom_point(aes(size= n)) + ggtitle("Emotions: Traditionalists") + facet_grid(~ emotion)

```

