---
title: "HW6"
author: "juan"
date: '2017-11-03'
output: html_document
---


```{r}
library(purrr)
library(stringr)
library(listviewer)
library(tibble)
library(png)
library(grid)
library(readxl)
suppressPackageStartupMessages(library(dplyr))
suppressMessages(library(gridExtra) )
suppressMessages(library(mgcv))
suppressMessages(library(jsonlite))
suppressPackageStartupMessages(library(tidyverse))
```

# Welcome to homework 6 part 1. 
In this file I'll work the exercises in the Strings chapter or R for Data Science.In part two, I'll work on Halloween Datset if you want to see it just chick here --> https://github.com/juansbr7/STAT547Bonilla_Juan/blob/master/HW6-2.Rmd.

# PART 1
**summary of functions:**
str_length(), str_c(), str_replace_na(), str_sub(), str_to_lower(), str_to_upper(), str_wrap(), str_trim(), str_view(), str_view_all()


**NOTE:** Code that return too many resuts is commented, so it won't work until uncommented, .md file was not possible to generate.

## 14.2.5 String basics Exercises: (short answers in the exercises, detailed answers in the code)

**A and B.** In code that doesn’t use stringr, you’ll often see paste() and paste0(). What’s the difference between the two functions? (**short answer=** no separators are set by defaul in paste0()) 
What stringr function are they equivalent to? (**short answer=** str_c()) 
How do the functions differ in their handling of NA? (**short answer=** str_c() allows to replace NA) 
In your own words, describe the difference between the sep and collapse arguments to str_c() (**short answer=** separators define the type separator between string, collapse concatenate all string into one). **More detailed answers are beside the code.**

```{r}
#prints strings indepedently and leaves a space separator by default
(p1<-paste(1:10, "cat", "dog", NA)) 

#concatenates all strings into one and leaves a space separator.
(p2<-paste(1:10, "cat","dog", NA, collapse = ",")) 

#defines the separator among strings and keeps strings independent
(p3<-paste(1:10, "cat", "dog", NA, sep= ","))

#defines sepatators and concatenate all strings into one.
(p4<-paste(1:10, "cat", "dog", NA, sep= ",", collapse = ","))
```

```{r}
# remove separators by default
(p01<-paste0(1:10, "cat", "dog", NA))

# concatenates all strings into one 
(p02<-paste0(1:10, "cat", "dog", NA, collapse= ","))

```

```{r}
# concatenates single strings but does not collapse them by default 
(strc1<- str_c(1:10, "cat", "dog")) 
#if there is a NA, it returns only NAs
(strc1<- str_c(1:10, "cat", "dog", NA))
# to print NA as NAs or to replace them, str_replace_na() is the way to go.
(strc2<- str_c(1:10, "cat", "dog", str_replace_na(NA, replacement = "-")))
```


C. Use str_length() and str_sub() to extract the middle character from a string. What will you do if the string has an even number of characters? 

```{r}
string1<- c("grapefruits", "apples", "lemons")
(singular<- str_sub(string1, 1,-2))

even<-"strawberry"
uneven<-  "raspberry"
str_sub(even,str_length(even)/2,str_length(even)/2+1)
str_sub(uneven,str_length(uneven)/2+0.5, str_length(uneven)/2+0.5)
```

D. What does str_wrap() do? When might you want to use it? Answer= format indentation of strings at the beggining and at the end of each line.  
```{r}
text<- c("string is character vector of strings to reformat. ", "width is positive integer giving target line width in characters. A width less than or equal to 1 will put each word on its own line.", "indent is non-negative integer giving indentation of first line in each paragraph ", "exdent is non-negative integer giving indentation of following lines in each paragraph")
str_wrap(text, width = 50, indent = 10, exdent = 0)
```

E. What does str_trim() do?  ANSWER= removes  whitespace in the string.
```{r}
string<- c("nospace","     l_&_r      ", "rightspace     ", "      leftspace")
(nospace<-str_trim(string))# removes all the whitespace
str_trim(string, "right")#removes right's whitespace
str_trim(string, "left")#removes left's whitespace
```

What’s the opposite of str_pad()? **short answer=** adds whitespace on the left or right until the string is the length indicated. If the string is shorter than indicated, it'll leave the string untouched, unless the str_trunc() is used. str_pad(number, with= lenght want it to be, side= where you want the padding, pad= what you want to)

```{r}
#adds right's whitespace until stings lenght 10
str_pad(nospace, 10, "right")
#adds left's 0 as padding until stings lenght 7
str_pad(nospace, 7, side="left", pad="0")
str_trunc(nospace, 8)
```

## 14.3.1.1 Matching patterns with regular expressions Exercises:

Explain why each of these strings don’t match a \: "\", "\\", "\\\". ANSWER= because R reads
x<- "a\\b"
str_view(x, "\")# # error, R thinks it is an escape of a regular expression
str_view(x, "\\") #  error, R thinks \\ refers to a regular expression.
str_view(x, "\\\")# error, R thinks it is an incomplete expression
```{r}
x<- "a\\b"
# one \ indicates an espape, the other two the creation of a regular expression and the one left is the character to be matched.
str_view(x, "\\\\") 

```

How would you match the sequence "'\
```{r}
(q<- "match this \"'\\ ")
str_view(q, "\"\'\\\\")
```



What patterns will the regular expression \..\..\.. match? How would you represent it as a string?

```{r}
rg<- "this homework is so l.o.n.g"
writeLines(rg)
str_view(rg, "\\..\\..\\..")


```

## 14.3.2 Anchors Exercises

How would you match the literal string "$^$"?
```{r}
(s<- "match this $^$")
str_view(s, "\\$\\^\\$")
```

Given the corpus of common words in stringr::words, create regular expressions that find all words that:

Since this list is long, you might want to use the match argument to str_view() to show only the matching or non-matching words. commented code is too long to print
```{r}
#Start with “y”
str_view(words,"^y", match = TRUE)
#End with “x”
str_view(words, "x$", match = TRUE) 
#str_view(words, "^...$", match = TRUE) #Are exactly three letters long. 
#str_view(words, ".......", match = TRUE) #Have seven letters or more.
```

## 14.3.3 Character classes and alternatives Exercises
Create regular expressions to find all words that:

```{r}
#str_view(words, "^[aeiou]", match = TRUE) #Start with a vowel.
str_view(words, "^[^aeiou]*$", match = TRUE)# no vowels anywhere, only consonants
str_view(words, "[^e]ed$", match = TRUE)# End with ed, but not with eed.
str_view(words, "(ing|ise)$", match = TRUE)# End with ing or ise.
```

Empirically verify the rule “i before e except after c”.
```{r}
str_view(words, "ei", match= TRUE)# words staring with ei and weight break the rule

```

Is “q” always followed by a “u”?
```{r}
str_view(words, "q[^u]", match= TRUE)# No results of words with q not followed by u
```

Write a regular expression that matches a word if it’s probably written in British English, not American English.
```{r}
str_view(words, "our$", match= TRUE )
```

Create a regular expression that will match telephone numbers as commonly written in your country.
```{r}
pn<- c("16137007177", "011573103835894", "011578245030") ###01157 is the area code of my country
str_view(pn, "^01157", match= TRUE )
```

## 14.3.4 Repetition Exercises

Describe the equivalents of ?, +, * in {m,n} form. 
  ? : {1}   0 or 1
\ * : {0, } 0 or more
 \+ : {1, } 1 or more 
```{r}
# REMINDER:
#{n}: exactly n 
#{n,}: n or more
#{,m}: at most m 
#{n,m}: between n and m 
#"\n" new line 
#"\t" tab 
#"\b" boundary 
#"\d" digit 
#"\s" whitespace
#[abc]: matches a, b, or c
#[^abc]: matches anything except a, b, or c
#^  $ : only match a complete string
#“.” matches any character including whitespace 
```
Describe in words what these regular expressions match: (read carefully to see if I’m using a regular expression or a string that defines a regular expression.)
^.*$ answer= 0 or 1 character(any)
"\\{.+\\}" answer= 1 or more of any character
\d{4}-\d{2}-\d{2} answer= 4digits-2digits-2digits
"\\\\{4}" answer= \ x 4 = \\\\

Create regular expressions to find all words that:
```{r}
str_view(words,"^[^aeiou]{3}", match= TRUE) #Start with three consonants.
str_view(words,"[aeiou]{3,}", match= TRUE) #Have three or more vowels in a row.
#str_view(words,"^[aeiou][^aeiou]{2,}$", match= TRUE) #Have two or more vowel-consonant pairs in a row.
```

## 14.3.5 Grouping and backreferences Exercises

Describe, in words, what these expressions will match:

(.)\1\1 3x1 defined character ex aaa
"(.)(.)\\2\\1" a and b characters, then b and a
(..)\1    1x 2 defined characters ex abab
"(.).\\1.\\1"  3x1 defined character with one random character inbetween ex afaja
"(.)(.)(.).*\\3\\2\\1" 3 defined characters, 0 or more random characters, 3 defined characters in reverse. ex a b c ytu c b a




```{r}

#str_view(words, "^(.).*\\1$", match = TRUE)    #Start and end with the same character.

str_view(words, "(..).*\\1", match = TRUE)#Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)

str_view(words, "(.).*\\1.*\\1", match = TRUE)#Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)

```



## 14.4.2 Detect matches Exercises
For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.
```{r}
str_view(words, "^x|x$", match= TRUE)
st<-str_detect(words,"^x")
end<-str_detect(words,"x$")
words[st | end]
```

Find all words that start or end with vowel and end with consonant.
```{r}
#str_view(words, "^[aeiou].*[^aeiou]$", match = TRUE)
vwl = str_detect(words, "^[aeiou]")
const = str_detect(words, "[^aeiou]$")
words[vwl & const]
```

Are there any words that contain at least one of each different vowel?
```{r}
a = str_detect(words, "a")
e = str_detect(words, "e")
i = str_detect(words, "i")
o = str_detect(words, "o")
u = str_detect(words, "u")
words[a & e & i & o & u ]
words[a & e & i & o ] # no luck with all the vowels, but some with no u
```


What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)
```{r}
(wordsdf<- tibble(word= words, vowel = str_count(word, "[aeiou]"), len= str_length(word), proportion= len/vowel)) %>% 
  arrange(desc(vowel))%>% head(10)

```


## 14.4.3 Extract matches Exercises

In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.
```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
truecolours <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")## add boundary at the begining and end and grouped them
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, truecolours, match = TRUE)
```

From the Harvard sentences data, extract:

```{r}
str_extract(sentences, "[^ ]+") %>% head(10)#The first word from each sentence.
p<-str_extract(sentences,"[^ ]+ing[ .]")#get words ending with ing
p[!is.na(p)] %>% str_sub(1, -2)%>% head(10)# removes NA and punctuation at the end
no_s<-str_extract(sentences,"[^ ]+s[ .]")  #end in s. Note= not all are nouns. no irregular plurals
no_s[!is.na(no_s)] %>% str_sub(1, -2)%>% head(10)
```

## 14.4.4 Grouped matches Exercises

Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.
```{r}
nbs<-"(one|two|three|four|five|six|seven|eight|nine|ten) ([^ ]+)"
both<- sentences %>% str_extract(nbs) 
both2<-both[!is.na(both)]
both2 %>% 
  str_match(nbs) %>% head(10)
```

Find all contractions. Separate out the pieces before and after the apostrophe.
```{r}
contraction<- "[^ ]+\\'[^ .]+"
contracted<-sentences %>% str_extract(contraction)
contracted[!is.na(contracted)] %>% head(10)
```
## 14.4.5 Replacing matches Exercises

Replace all forward slashes in a string with backslashes.
```{r}
backslash<-str_replace_all("this / need to change","/","\\\\")
writeLines(backslash)
```

Implement a simple version of str_to_lower() using replace_all().
```{r}
upstr<- "ThIs strIng Is crAzy"
str_replace_all(upstr,c("I" = "i", "A" = "a"))

```

Switch the first and last letters in words. Which of those strings are still words?
```{r}
backwards<- str_replace(words,"(^.)(.*)(.$)", "\\3\\2\\1")
intersect(words, backwards)


```

## 14.4.6 Splitting Exercise
REMINDER=  boundary(type = c("character", "line_break", "sentence", "word"), skip_word_none = NA, ...)
Split up a string like "apples, pears, and bananas" into individual components.
Why is it better to split up by boundary("word") than " "? boundary doesn't require to specify puntuation or whitespace
```{r}
example<- c("apples, pears, and bananas")
str_split(example,", and |," )[[1]]
str_split(example, boundary("word") )[[1]]
```


What does splitting with an empty string ("") do? Experiment, and then read the documentation. split by characters
```{r}
str_split(example,"" )
```
## 14.5.1 Exercises

How would you find all strings containing \ with regex() vs. with fixed()?

```{r}
fixedvsregex<-c("there are many \\\\\\\ here", "and \\\\\\\\ here")
str_view_all(fixedvsregex, fixed("\\\\"))
str_view_all(fixedvsregex, "\\\\")
```
What are the five most common words in sentences?
```{r}
w <-str_split(sentences, boundary("word"), simplify = TRUE) %>% as.vector()
tibble(wordsinsentence = str_to_lower(w)) %>%
  group_by(wordsinsentence) %>%
  summarise(counts = n()) %>%
  arrange(desc(counts)) %>%
  head(6) %>%
  tail(5)
```



```{r}

```


