---
title: "HMWK2"
author: "juan"
date: '2017-09-26'
output: html_document
---
H1 Welcome to Juan's Homework 2
In this assignment, I'm going to show the increment of population, life expectancy and gdpper capita in Colombia and other contries in the Americas since the year I was born (**1987**). This is to prove my theory that I brought good things to the world.

```{r}
library(gapminder)
library(tidyverse)

typeof(gapminder)
a<- filter(gapminder, continent == "Americas")

select(filter(a, country== "Colombia"), country, gdpPercap, lifeExp, year) 

```
Here we have Life Expectancy, gdpPercapita and Population in Colombia since 1987 to 2004; followed by graph.
```{r}
my_colombia<- a %>% 
  filter(country== "Colombia", year >= 1987) %>%
  select(country, year, lifeExp, gdpPercap, pop)
my_colombia
```
```{r pressure, echo=FALSE}
ggplot(my_colombia, aes(year, lifeExp))+ geom_path()
```
As we can see in the graph there is a significant improvement of life expectancy. Here are the increment numbers and the function that I wrote to see the results:
```{r}
my_colombia
inc<- function(x,y) {z<- x - y}
mutate(my_colombia, inc_lifeExp= inc(lifeExp, lifeExp[+1]), inc_gdpPercap= inc(gdpPercap, gdpPercap[+1]))
```
Now, let's see the population vs gdpPercap 
```{r}
my_colombia %>%
  arrange(year) %>% 
  ggplot(aes(pop, gdpPercap)) +
  geom_point(color= "red") +
  geom_path(color= "blue")
```
But enough about Colombia, let's see how other countries in the Americas have done since I came to earth:
```{r}
a %>%
  group_by(country) %>% 
  ggplot(aes(year, gdpPercap)) + geom_smooth()+
  geom_point(aes(color= country, size= pop ))
```


