---
title: "HW#"
author: "juan"
date: '2017-09-27'
output: github_document
---

```{r}
library(gapminder)
library(tidyverse)
```

## Homework 3 --work in progress--
+ Get the maximum and minimum of GDP per capita for all continents.
Look at the spread of GDP per capita within the continents.
+ Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.

+ How is life expectancy changing over time on different continents?

+ Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy – you decide – a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year.

+ Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.

+ Make up your own! Between the dplyr coverage in class and the list above, I think you get the idea.


```{r}

```

##  Plots



```{r}

```


