---
title: "HW5"
author: "juan"
date: '2017-10-18'
output: github_document
---
#####1. GAPMINDER OPTION
+ Drop Oceania. Filter the Gapminder data to remove observations associated with the continent of Oceania. Additionally, remove unused factor levels. Provide concrete information on the data before and after removing these rows and Oceania; address the number of rows and the levels of the affected factors.

```{r}
library(gapminder)
suppressMessages(library(tidyverse))
library(forcats)
library(ggplot2)
suppressMessages(library(gridExtra) )
suppressMessages(library(mgcv))
library(png)
library(grid)
library(RColorBrewer)

```

+ Let's start by exploring the factors in the gapminder dataframe. There are 5 levels in the continent factor and 142 levels in the countries, the number of rows (obs) 1704. So if we drop Ocenia and check the levels we should expect to see 4 levels. Oceania only has two countries in this dataframe so we should see 140 levels in the country factor and less than 1704 rows.
```{r}
str(gapminder)
gapminder %>% filter(continent== "Oceania") %>% count(country)
conts <- c("Americas", "Europe", "Asia", "Africa")
no_oceania <- gapminder %>%
  filter(continent %in% conts)
str(no_oceania)
```


+ Ooops The number of rows decreased to 1680 but there are still 5 levels and 142 levels which means that the filter did not dropped the factot levels belonging to country and continent.  Let's use  droplevels().

```{r}
no_oceania <- gapminder %>%
  filter(continent %in% conts) %>% 
  droplevels() 

str(no_oceania)
nrow(no_oceania)
```

Ok, dropped. Now we really have 4 continents. Now, I'll order the countries based on their population, from the largest to the lowest. It is a pretty long list so I'll disploy only the first 5. 
```{r}
fct_reorder(no_oceania$country, no_oceania$pop, max, .desc = TRUE) %>% levels() %>% head()
  
```

+ Does merely arranging the data have any effect on, say, a figure? what effect does this have on a figure?
Let's see figures with ordered and unordered levels. I'll save my graphs in ***.png*** format using ggsave and I'll use ***saveRDS()*** to save my data. 
```{r figure 1}
americas_data <- gapminder %>% filter( continent == "Americas") %>% mutate(pop= pop/10^6) 
plot_a<- ggplot(americas_data, aes(x = pop, y = country)) + geom_point(aes(colour= year))+ labs(x="", y="")+ guides(colour= FALSE)
plot_b<- ggplot(americas_data, aes(x = pop, y = fct_reorder(country, pop))) +
  geom_point(aes(colour= year)) + labs(x="", y="")+ 
  theme( legend.position = c(.95, .10),
  legend.justification = c("right", "bottom"),
  legend.box.just = "right",
  legend.margin = margin(6, 6, 6, 6))
  
grid.arrange(plot_a, plot_b, ncol= 2, top= "Levels= unordered vs ordered", left= "countries", bottom= "Population in Millions")
saveRDS(americas_data, "americas_data.rds")
ggsave("plot_a.png", plot_a)
ggsave("plot_b.png", plot_b)
```

#####Visualization design: 
+ Here I will work with the  mean of the gdpPercapita and ordered values according to bothe the max and min gdpPercapita in the Americas. In this case I'll save my graphs independently in .png and the grids as pdfs. I'll also toy a bit more with the background, the shapes and colors in the graphs.

```{r figure 2}
img<-readPNG("dollar.png")
g <- rasterGrob(img, interpolate=TRUE)
ordered_americas_data <- readRDS("americas_data.rds") %>% droplevels() %>% group_by(country)  %>%   summarise(meangdp= mean(gdpPercap), maxgdp = max(gdpPercap), mingdp = min(gdpPercap)) %>% mutate(country = fct_reorder(country, mingdp, min))
ordered_americas_data %>% head(10) %>% knitr::kable(format = "markdown", padding=2, caption= "default alphabetical order")
levels(ordered_americas_data$country) %>% head(10) %>%   knitr::kable(format = "markdown", padding=2, col.names =  "reordered based on mingdp")
plot_c<- ggplot(ordered_americas_data, aes(x = meangdp, y = country)) + geom_point(aes(mingdp), colour= "orangered4", size= 5, shape= "-")+ geom_point(aes(maxgdp), colour= "darkgreen", size= 5, shape= "+")+ labs(y= "", x= "")+ ggtitle(" Ordered by Mingdp(-)")+ annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)

plot_d<- ggplot(ordered_americas_data, aes(x = meangdp, y = fct_reorder(country, maxgdp))) + geom_point(aes(mingdp), colour= "orangered4", size= 5, shape= "-")+ geom_point(aes(maxgdp), colour= "darkgreen", size= 5, shape= "+")+ labs(y= "", x= "")+ ggtitle("Ordered by Maxgdp(+)")+ annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)
ggsave("plot_b.png", plot_b)
grid_a_b<- grid.arrange(plot_c, plot_d, ncol= 2, top= "GdpPercapita in America ordered by the Maximun and the Minimun", left= "countries", bottom= "gdpPercapita")
grid_a_b
pdf("grid_a_b.pdf", width=12, height=6)


```

+ Here I took two of the graphs from previous assignements and tweaked them based on recommendations of the peer reviewers and TAs. I also gave color brewer a try and used some of their palettes (Set3).

```{r figure 3}
 south_america<-  c("Colombia", "Chile", "Argentina", "Uruguay", "Ecuador", "Bolivia", "Paraguay", "Peru", "Brazil", "Venezuela")

s_america<- gapminder %>% filter(country %in% south_america) %>% droplevels() %>%  group_by(country)
write.csv(s_america, "s_america.csv") 
 
plot_c<-ggplot(s_america,aes(year, gdpPercap)) +
  geom_point(aes(color= country, size= pop/10^6 ))+ geom_line(aes(colour= country))+ scale_size_continuous("Population in Millions") + scale_colour_manual(values=brewer.pal(n=10, "Set3"))
plot_c
ggsave("plot_c.png", plot_c)
```
+ Now let's look at some min, max life expectancies in all continents.

```{r figure 4}
lifeEx_world<- gapminder %>%
  group_by(year, continent) %>% 
  select(lifeExp, continent) %>% 
summarise(mean_le= mean(lifeExp, trim= 0.2), max_le= max(lifeExp), min_le= min(lifeExp), md_le= median(lifeExp)) 

maxlifeEx_world<-lifeEx_world %>% 
  select(year, max_le, continent) %>% 
    spread(key = year, value = max_le) %>% 
  knitr::kable(format = "markdown", padding=2, caption= "Max Life Expectancy per year in all Countries")
maxlifeEx_world
minlifeEx_world<-lifeEx_world %>% 
  select(year, min_le, continent) %>% 
    spread(key = year, value = min_le) %>% 
  knitr::kable(format = "markdown", padding=2, caption= "Min Life Expectancy per year in all Countries")
minlifeEx_world
plot_d<- lifeEx_world %>% 
  ggplot(aes(x= year, y= mean_le )) +
  facet_grid(~ continent)+
    geom_ribbon(aes(ymin= min_le, ymax= max_le), fill= "darkblue")+
  geom_line(aes(y= mean_le, colour= "mean_le"))+
      theme(axis.text.x = element_text(angle=90),
          plot.title = element_text(hjust=0.5))+
    geom_line(aes(y=md_le , colour= "md_le"))+
      ggtitle("Max, Min, Mean and Median Life Expectancy in all Continents")+
      ylab("Life Expectancy")+
      xlab("years")
plot_d
ggsave("plot_d.png", plot_d)
```
 + Extra bit: I guess gapminder forgot to add the important detail of weird_food in some countries. We all know that Japan is known for its sushi, but also, for its tuna eyeballs. Let's see others:
 
```{r figure 5}
c<-  c("France", "Thailand", "United States", "China", "Japan", "Ecuador", "Cambodia", "Philippines", "Mexico", "Australia")
food<- c("frog legs", "fried insects", "fried rattlesnake", "rotten eggs", "tuna eyeballs", "guinea pig", "fried spider", "fertilized egg", "insect eggs", "grubs")
country<- factor(c)
weird_food<-factor(food)
newdf<- data_frame(country, weird_food)
newdf
dataset1<- gapminder %>%
  filter(country %in% c, year== 2007) %>%
           droplevels() 
both<-left_join(dataset1, newdf, by= "country")
both%>% knitr::kable()

ggplot(both, aes(country, weird_food, fill= weird_food))+ geom_bar(stat= "identity", colour= "black")+ scale_fill_manual(values=brewer.pal(n=10, "Paired"))+ labs(x="", y= "") +
    theme(axis.text.x = element_text(angle=90))+ ggtitle("Weird, Gross and Tasty")

```

###### Reflection

I had so much fun tweaking graphs and thinking about new things to do. Thanks 





